---
title: "p8105_hw1_vy2196"
output: github_document

---

### Problem 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(moderndive)   # load the package
data("early_january_weather")  # load dataset
head(early_january_weather) #first few rows and columns of data
```

#### Data description
The `early_january_weather` dataset contains `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. 

It includes hourly weather observations from New York City in early January 2013. 
Key variables include:

- `time_hour`: the date-time of observation  
- `temp`: temperature in Fahrenheit  
- `humid`: relative humidity (0–100)  
- `wind_dir`: wind direction
- `wind_speed`: wind speed in mph  
- `visib`: visibility in miles  
- plus other variables describing conditions such as `pressure`and `origin`.

The mean temperature during this period was 
`r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F.

#### Creating a scatterplot

```{r scatterplot, fig.width=7, fig.height=4}
library(ggplot2)

p <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7) +
  scale_color_viridis_c(option = "plasma") +
  labs(title = "NYC Temperature in Early January 2013",
       x = "Time (Hour)",
       y = "Temperature (°F)",
       color = "Humidity") +
  theme_minimal()

# Show in knitted report
p

# Exporting scatterplot
ggsave("scatterplot_temp_vs_time.png", plot = p, width = 7, height = 4)
```

##### Descriptions of the graph's pattern
The scatter plot shows a gradual increase in temperature over the next two weeks starting from January 1st. There was a big temperature difference in the beginning of January and then it slowly decreased and increased by mid-January. The humidity also increased comparing to early January from 40s to approaching 100 toward mid-January. 

### Problem 2

```{r}
set.seed(123)  # for reproducibility
num_var <- rnorm(10)
log_var <- num_var > 0
char_var <- c("a","b","c","d","e","f","g","h","i","j")
fact_var <- factor(rep(c("low", "medium", "high"), length.out = 10))
df <- data.frame(num_var, log_var, char_var, fact_var)
df
```


